A Pandemic-Scale Ancestral Recombination Graph for SARS-CoV-2

### 2025-10-16 latex/main.tex
Updates from Overleaf

\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{recombinant_evidence.pdf}
\caption{
Properties of recombination events. A,B,C 
show summaries of the 354 sc2ts recombination nodes
passing quality control filters
colored by 3SEQ classification (recombinant: purple, non-recombinant: gray).
(A) Upper: jittered scatterplot comparing number of mutations
when recombination allowed ($y$: right axis)
versus an alternative copying path disallowing recombination ($x$: top axis);
recombination nodes in Table~\ref{tab:pango_x_lineages} are labeled starting with `X' and
those identified by Jackson et al.\cite{Jackson2021} are prefixed with a `J', e.g.
Jackson group B (JB) and Jackson singleton recombinant 1 (J1).
Point sizes reflect number of descendants.
Lower: recombination nodes classified by number of mutations averted
by allowing recombination (i.e. $x-y$).
(B) Distribution of time to the MRCA of the parents of each QC-passing recombination node.
(C) Distribution of the Pango lineage distance between parents.
(D) Copying patterns (STAR Methods, Figure~\ref{fig:copying_pattern_examples}) 
for XA, XW, XBB, and an
undesignated recombination event under XBB, which we refer to as XBB.x.}
\label{fig:recombinant_evidence}
\end{figure}

Figure~\ref{fig:recombinant_evidence}A plots
the number of mutations required
in the recombinant versus non-recombinant
Viterbi solutions (upper scatter plot)
and the resulting distribution of the number of
averted mutations (lower histogram)
over all 354 recombination events that pass QC.
% The 5\% of QC-passing recombinants
% not validated by 3SEQ are also among the most marginally supported
% (see next section; Figure~\ref{fig:recombinant_evidence}).
Also shown is the 3SEQ classification status of these recombinants.
All of the 12\% of recombinants that do not pass 3SEQ validation are
from the most weakly supported, with 4 or 5 averted mutations.
Figures~\ref{fig:recombinant_evidence_cov_recomb} and
~\ref{fig:recombinant_evidence_rebar} 
plot the same information for CovRecomb and rebar validation,
showing that their validation status tends to be function
of the divergence between the recombinant parents.

>>>>

\begin{figure}[h]
\centering
\includegraphics[width=1\linewidth]{recombinant_evidence.pdf}
\caption{
Properties of recombination events. A,B,C 
show summaries of the 354 sc2ts recombination nodes
passing quality control filters
colored by 3SEQ classification (recombinant: purple, non-recombinant: gray).
(A) Upper: jittered scatterplot comparing number of mutations
when recombination allowed ($y$: right axis)
versus an alternative copying path disallowing recombination ($x$: top axis);
recombination nodes in Table~\ref{tab:pango_x_lineages} are labeled starting with `X' and
those identified by Jackson et al.\cite{Jackson2021} are prefixed with a `J', e.g.
Jackson group B (JB) and Jackson singleton recombinant 1 (J1).
Point sizes reflect number of descendants.
Lower: recombination nodes classified by number of mutations averted
by allowing recombination (i.e. $x-y$).
(B) Distribution of time to the MRCA of the parents of each QC-passing recombination node.
(C) Distribution of the Pango lineage distance between parents.
(D) Copying patterns (STAR Methods, Figure~\ref{fig:copying_pattern_examples}) 
for XA, XW and XBB.}
\label{fig:recombinant_evidence}
\end{figure}

Figure~\ref{fig:recombinant_evidence}A plots
the number of mutations required
in the recombinant versus non-recombinant
Viterbi solutions (upper scatter plot)
and the resulting distribution of the number of
averted mutations (lower histogram)
over all 354 recombination events that pass QC.
Also shown is the 3SEQ classification status of these recombinants.
The 5\% of recombinants that do not pass 3SEQ validation are
among the most weakly supported, with 4 or 5 averted mutations.
[TODO discuss B and C]
Figures~\ref{fig:recombinant_evidence_cov_recomb} and
~\ref{fig:recombinant_evidence_rebar} 
plot the same information for CovRecomb and rebar validation,
showing that their validation status tends to be function
of the divergence between the recombinant parents.
